---
title: MEPS
output: html_notebook
---

Set up packages
```{r}
install.packages("tidyverse")
install.packages("dplyr")
install.packages("tibble")
install.packages("foreign")  # Only need to run these once
 install.packages("survey")

 library(foreign) # Run these every time you re-start R
 library(survey)
 library(tidyverse)
 library (dplyr)
 
 setwd("C:/Users/GrossKennen/OneDrive for Business/MEPS/MEPS_R")
 library(readxl)
CondCodes <- read_excel("C:/Users/GrossKennen/OneDrive for Business/MEPS/CODES/MEPSConditionCodes.xlsx")
View(CondCodes)
 
```

Load IP data and MEPS CCS Service Line grouper
```{r}
library(haven)
IP2015 <- read_sav("C:/Users/GrossKennen/OneDrive for Business/MEPS/2015/IP/IP2015.sav")
View(IP2015)

library(readxl)
CondCodes <- read_excel("C:/Users/GrossKennen/OneDrive for Business/MEPS/CODES/MEPSConditionCodes.xlsx")
View(CondCodes)
```

Join IP 2015 to master grouper of CCS to SIngle Tier Service Line 
```{r}



IP2015 <- transform(IP2015, CCS = as.numeric(IPCCC1X))
CondCodes$CCS<-as.numeric(CondCodes$CCSCodes)
IP2015 <-IP2015 %>% left_join (CondCodes, c("CCS" = "CCS"))
```

Turn specialty tier column into one column per tier with 1, 0 
```{r}

rename(IP2015, tier = SingleTierServiceLine)
IP15_spread <- IP2015 %>% mutate (yesno=1) %>% distinct %>% spread (SingleTierServiceLine, yesno, fill = 0) 
```

Load Population file
```{r}
library(haven)
pop2015 <- read_sav("C:/Users/GrossKennen/OneDrive for Business/MEPS/2015/Population/Population2015.sav")

```
Recode Age
```{r}
pop2015<- mutate (pop2015, agecat4 = ifelse (AGELAST <18, "0-17", ifelse (AGELAST >17 & AGELAST<45, "18_44", ifelse (AGELAST >44 & AGELAST<65, "45-64", "65+"))))
pop2015<- mutate (pop2015, agecat4num = ifelse (AGELAST <18, 1, ifelse (AGELAST >17 & AGELAST<45, 2, ifelse (AGELAST >44 & AGELAST<65, 3, 4))))


```
Recode Insurance 
```{r}
pop2015<- mutate (pop2015, ins5_num = ifelse (INSURC15 == 1, 1, ifelse (INSURC15 == 2, 2, ifelse (INSURC15 ==4 | INSURC15 ==5, 3, ifelse (INSURC15 == 7 | INSURC15 == 3, 4, ifelse (INSURC15 == 6, 5, ifelse (INSURC15 == 8, 3,0)))))))

pop2015<- mutate (pop2015, ins5= ifelse (INSURC15 == 1, "Commercial", ifelse (INSURC15 == 2, "Medicaid", ifelse (INSURC15 ==4 | INSURC15 ==5, "Medicare", ifelse (INSURC15 == 7 | INSURC15 == 3, "Uninsured", ifelse (INSURC15 == 6, "Dual", ifelse (INSURC15 == 8, "Medicare",0)))))))
```

Create Table of Freq of Tier 1 by unique ID
```{r}
IP15count <- select (IP15_spread, DUPERSID, Cardiovascular: Urology) %>% group_by (DUPERSID ) %>% summarize (Cardiovascular_sum = sum(Cardiovascular), 
Digestive_sum = sum(Digestive),
ENT_sum = sum(ENT),
Gynecology_sum = sum(Gynecology),
Hepatobiliary_sum = sum(Hepatobiliary),
Infection_sum = sum(Infection),
Neuroscience_sum = sum(Neuroscience),
Obsetetrics_sum = sum(Obsetetrics),
Oncology_sum = sum(Oncology),
Ophthalmology_sum = sum(Ophthalmology),
Orthopedics_sum = sum(Orthopedics),
OtherMedical_sum = sum(OtherMedical),
Psychiatry_sum = sum(Psychiatry),
Respiratory_sum = sum(Respiratory),
Skin_sum = sum(Skin),
TraumaInjuries_sum = sum(TraumaInjuries),
Urology_sum = sum(Urology),
Endocrinology_sum = sum(Endocrinology),
Hematology_sum = sum(Hematology))

```

Join IP counts by Service to Pop file 
```{r}

pop2015small <- pop2015 %>% select(DUPERSID, ins5, agecat4, agecat4num, VARPSU, VARSTR,PERWT15F)

pop2015small <- transform(pop2015small, DUPERSID = as.numeric(DUPERSID))
IP15count <- transform(IP15count, DUPERSID = as.numeric(DUPERSID))
pop2015small <-pop2015small %>% left_join (IP15count, c("DUPERSID" = "DUPERSID"))

pop2015small[is.na(pop2015small)]<- 0

variable.names(pop2015small)

```


```{r}

rm(CondCodes, IP15_group, IP15_spread, IP15count, IP2015, MEPSConditionCodes, pop2015)

options(survey.lonely.psu='adjust')

mepsdsgn = svydesign(id = ~VARPSU,
                     strata = ~VARSTR,
                     weights = ~PERWT15F,
                     data = pop2015small,
                     nest = TRUE)  





MepsIPRates15 <- svyby(~Cardiovascular_sum:~Gynecology_sum:~Oncology_sum:~Infection_sum:~ENT_sum:~Hepatobiliary_sum:~Neuroscience_sum:~Obsetetrics_sum:~Ophthalmology_sum:~Orthopedics_sum: ~OtherMedical_sum:~Psychiatry_sum:~Respiratory_sum:~Skin_sum:~TraumaInjuries_sum:~Urology_sum:~Endocrinology_sum:~Hematology_sum: ~Digestive_sum , ~agecat4:~ins5, mepsdsgn, svymean)

MepsIPRates15<- select(MepsIPRates15, -ins5)

library(foreign)
write.foreign(as.data.frame(MepsIPRates15), "C:/Users/GrossKennen/OneDrive for Business/iVantage/Forecasting/Data4/exported_MepsIPRate15b.txt", "C:/Users/GrossKennen/OneDrive for Business/iVantage/Forecasting/Data4/exported_MepsIPRate15b.sps", package="SPSS")

              
            

